Bayesian Hyper-Heuristics

### 2023-02-04 article/frontmatter/abstract.tex
✨ (article): Introduction Done
🚧 (article): Save point, WIP
🚧 (article): WIP save point for revision

\begin{abstract}

	This work introduces a novel population-based \Acf{BHH}. An empirical study is done by using the \acs{BHH} to train \acsp{FFNN}. An in-depth behaviour analysis is done and the performance of the \acs{BHH} is compared to that of ten popular low-level heuristics, each with different search behaviours. The chosen heuristic pool consists of classic gradient-based heuristics as well as meta-heuristics. The empirical process is executed on fourteen datasets consisting of classification and regression problems with varying characteristics. The \acs{BHH} is shown to be able to train \acsp{FFNN} well and provide an automated method for finding the best heuristic to train the \acsp{FFNN} at various stages of the training process.

\end{abstract}

>>>>

\begin{abstract}
	This work introduces a novel population-based \acf{BHH} that is used to train \acfp{FFNN}. An in-depth behaviour analysis is done and the performance of the \acs{BHH} is compared to that of ten popular low-level heuristics, each with different search behaviours. The chosen heuristic pool consists of classic gradient-based heuristics as well as \acf{MH}. The empirical process is executed on fourteen datasets consisting of classification and regression problems with varying characteristics. The \acs{BHH} is shown to be able to train \acp{FFNN} well and provide an automated method for finding the best heuristic to train the \acp{FFNN} at various stages of the training process.
\end{abstract}

### 2023-02-04 article/frontmatter/highlights.tex
🚧 (article): WIP save point for revision

%%Research highlights
\begin{highlights}
	\item A novel heuristic selection operator is used that focuses on using Bayesian statistics to calculate the probability that a heuristic should be selected in order to efficiently train \acp{FFNN}. The resulting \acs{HH} is referred to as the \acf{BHH}.

	\item The results of the empirical study show with statistical significance and certainty that the \Acs{BHH} performs generally well on multiple problems. It is shown that, for each problem, the \Acs{BHH} performance is comparable to the best low-level heuristics included in the heuristic selection pool.

	\item The results of the empirical study show that the \Acs{BHH} is able to select the best heuristic to train \acp{FFNN} in general. This relieves researchers from the burden of having to do this selection process manually through trial and error.

	\item The results of the empirical study show that the \Acs{BHH}, given a diverse set of lower-level heuristics, will generally produce good results when applied to multiple problems at the same time.

	\item Finally, the results of the empirical study show that the \Acs{BHH} is capable of utilising \textit{a priori} \footnote{Latin word, meaning ``from what comes before''.} knowledge in which a predefined selection bias is used for heuristics that are known to be well suited for certain problems.
\end{highlights}

>>>>

\begin{highlights}
	\item A novel \acf{HH} selection operator is used that focuses on using Bayesian statistics to calculate the probability that a heuristic should be selected in order to efficiently train \acfp{FFNN}. The resulting \acs{HH} is referred to as the \acf{BHH}.

	\item The results of the empirical study show with statistical significance and certainty that the \acs{BHH} performs generally well on multiple problems. It is shown that, for each problem, the \acs{BHH} performance is comparable to the best low-level heuristics included in the heuristic selection pool.

	\item The results of the empirical study show that the \acs{BHH} is able to select the best heuristic to train \acp{FFNN} in general. This relieves researchers from the burden of having to do this selection process manually through trial and error.

	\item The results of the empirical study show that the \acs{BHH}, given a diverse set of lower-level heuristics, will generally produce good results when applied to multiple problems at the same time.

	\item Finally, the results of the empirical study show that the \acs{BHH} is capable of utilising \textit{a priori} \footnote{Latin word, meaning ``from what comes before''.} knowledge in which a predefined selection bias is used for heuristics that are known to be well suited for certain problems.
\end{highlights}

### 2023-02-04 article/sections/01-introduction.tex
🚧 (article): WIP save point for revision

A popular field of focus for studying \acp{ANN} is the process by which these models are trained. \acp{ANN} are trained by optimisation algorithms known as heuristics Many different heuristics have been developed and used to train \acp{ANN}  \cite{ref:gudise:2003, ref:rakitianskaia:2012, ref:montana:1989}. Each of these heuristics have different search behaviours, characteristics, strengths and weaknesses. Finding the best heuristic to train the \acs{ANN} is required in order to yield optimal results. This process is often non-trivial and could be a time-consuming exercise. Selection of the best heuristic to train \acp{ANN}, is often problem specific \cite{ref:allen:1996, ref:drake:2020, ref:pillay:2018}.

@@

This research takes a particular interest in developing a population-based, selection \acs{HH} that makes use of probability theory and Bayesian statistical concepts to guide the heuristic selection process. This research presents the novel \Acf{BHH}, a new high-level heuristic that utilises a statistical approach, referred to as \textit{Bayesian analysis}, which combines prior information with new evidence to the parameters of a selection probability distribution. This selection probability distribution is the mechanism by which the \acs{HH} selects appropriate heuristics to train \acp{FFNN} during the training process.

The remainder of this article is structured as follows: Section  \ref{sec:anns} provides background information on \acp{ANN}. Section  \ref{sec:heuristics} provides details on various types of heuristics that have been used to train \acp{FFNN}. Section  \ref{sec:hhs} presents background information on \acp{HH} and meta-learning. Section  \ref{sec:probability} presents background information on probability theory. Section  \ref{sec:bhh} presents the developed \Acs{BHH}. Section  \ref{sec:methodology} presents a detailed description of the empirical process and the setup of each experiment. Section  \ref{sec:results} provides and discusses the results of the empirical study. Section  \ref{sec:conclusion} summarises the research that is done along with a brief overview of the findings made throughout the research process.

>>>>

A popular field of focus for studying \acfp{ANN} is the process by which these models are trained. \acp{ANN} are trained by optimisation algorithms known as heuristics. Many different heuristics have been developed and used to train \acp{ANN} \cite{ref:gudise:2003, ref:rakitianskaia:2012, ref:montana:1989}. Each of these heuristics have different search behaviours, characteristics, strengths and weaknesses. Finding the best heuristic to train \acp{ANN} is required in order to yield optimal results. This process is often non-trivial and could be a time-consuming exercise. Selection of the best heuristic to train \acp{ANN}, is often problem specific \cite{ref:allen:1996, ref:drake:2020, ref:pillay:2018}.

@@

This research takes a particular interest in developing a population-based, selection \acs{HH} that makes use of probability theory and Bayesian statistical concepts to guide the heuristic selection process. This research presents a novel \acf{BHH}, a new high-level heuristic that utilises a statistical approach, referred to as \textit{Bayesian analysis}, which combines prior information with new evidence to the parameters of a selection probability distribution. This selection probability distribution is the mechanism by which the \acs{HH} selects appropriate heuristics to train \acfp{FFNN} during the training process.

The remainder of this article is structured as follows: Section \ref{sec:anns} provides background information on \acp{ANN}. Section \ref{sec:heuristics} provides details on various types of heuristics that have been used to train \acp{FFNN}. Section \ref{sec:hhs} presents background information on \acp{HH} and meta-learning. Section \ref{sec:probability} presents background information on probability theory. Section \ref{sec:bhh} presents the developed \acs{BHH}. Section \ref{sec:methodology} presents a detailed description of the empirical process and the setup of each experiment. Section \ref{sec:results} provides and discusses the results of the empirical study. Section \ref{sec:conclusion} summarises the research that is done along with a brief overview of the findings made throughout the research process.

### 2023-02-04 article/sections/02-artificial-neural-networks.tex
🚧 (article): WIP save point for revision

\acp{ANN} can generally be described as well-organised structures of mathematical computation and are inspired by the biological brain  \cite{ref:engelbrecht:2007}. \acp{ANN} have been successfully applied to a range of problem classes.  \citeauthor{ref:engelbrecht:2007}  \cite{ref:engelbrecht:2007} summarises some common problems that are solved using \acp{ANN}. These include classification  \cite{ref:khan:2001}, pattern matching  \cite{ref:cannady:1998, ref:kumar:1994}, pattern completion  \cite{ref:dayhoff:2001}, optimisation  \cite{ref:specht:1991} and data mining  \cite{ref:singh:2009}. \Acp{AN} provide the fundamental building blocks for \acp{ANN}.

The fundamental building blocks of \acp{ANN} is the \acs{AN}. The \acs{AN} implements a non-linear mapping from some $I$-dimensional, real-numbered input space to some $T$-dimensional, real-numbered output space. The output space is usually in the ranges $[0,1]$ or $[-1,1]$, depending on the problem being solved \cite{ref:engelbrecht:2007}. In essence, the \acs{AN} implements a weighted sum of its inputs and applies some non-linear transformation as a decision making boundary.

Multiple \acp{AN} can be organised and used together forming a ``network'' of \acp{AN} referred to as an \acf{ANN}. The architecture of the \acs{ANN} refers to the way in which \acp{AN} are organised. \acp{ANN} can be organised in layers where a single layer can contain multiple \acp{AN}. Output from one layer is propagated as input to the next layer. The topology of the \acs{ANN} refers to the way that layers of \acp{AN} are connected to each other. There are many different topologies  \cite{ref:miikkulainen:2010}.

This research focuses on a particular type of \acs{ANN}, referred to as \acp{FFNN}. \acp{FFNN} were the first and simplest type of \acp{ANN} developed  \cite{ref:schmidhuber:2015} and implement an architecture consisting of input, hidden and output layers by arranging them in sequential order. Furthermore, \acp{FFNN} implement fully connected topologies, where each \acs{AN} in one layer is connected to all \acp{AN} in the next, without any cycles  \cite{ref:zell:1994}.In \acp{FFNN}, information moves forward, in one direction, from the input nodes, through the hidden nodes and finally to the output nodes.

>>>>

\acp{ANN} can generally be described as well-organised structures of mathematical computation and are inspired by the biological brain \cite{ref:engelbrecht:2007}. \acp{ANN} have been successfully applied to a range of problem classes. \citeauthor{ref:engelbrecht:2007} \cite{ref:engelbrecht:2007} summarises some common problems that are solved using \acp{ANN}. These include classification \cite{ref:khan:2001}, pattern matching \cite{ref:cannady:1998, ref:kumar:1994}, pattern completion \cite{ref:dayhoff:2001}, optimisation \cite{ref:specht:1991} and data mining \cite{ref:singh:2009}. \acfp{AN} provide the fundamental building blocks for \acp{ANN}.

The \acs{AN} implements a non-linear mapping from some $I$-dimensional, real-numbered input space to some $T$-dimensional, real-numbered output space. The output space is usually in the ranges $[0,1]$ or $[-1,1]$, depending on the problem being solved \cite{ref:engelbrecht:2007}. In essence, the \acs{AN} implements a weighted sum of its inputs and applies some non-linear transformation as a decision making boundary.

Multiple \acp{AN} can be organised and used together forming a ``network'' of \acp{AN}, referred to as an \acs{ANN}. The architecture of the \acs{ANN} refers to the way in which \acp{AN} are organised. \acp{ANN} can be organised in layers where a single layer can contain multiple \acp{AN}. Output from one layer is propagated as input to the next layer. The topology of the \acs{ANN} refers to the way that layers of \acp{AN} are connected to each other. There are many different topologies \cite{ref:miikkulainen:2010}.

This research focuses on a particular type of \acs{ANN}, referred to as \acfp{FFNN}. \acp{FFNN} were the first and simplest type of \acp{ANN} developed \cite{ref:schmidhuber:2015} and implement an architecture consisting of input, hidden and output layers by arranging them in sequential order. Furthermore, \acp{FFNN} implement fully connected topologies, where each \acs{AN} in one layer is connected to all the \acp{AN} in the next, without any cycles \cite{ref:zell:1994}. In \acp{FFNN}, information moves forward, in one direction, from the input nodes, through the hidden nodes and finally to the output nodes.

### 2023-02-04 article/sections/03-heuristics.tex
🚧 (article): WIP save point for revision

Gradient-based heuristics are optimisation techniques that make use of derivates obtained from evaluating the \acs{ANN} being trained. In the context of supervised learning, loss functions produce a scalar value $\epsilon$, that represents the error between the output of the \acs{ANN} and the desired output. When using \acs{GD} to train \acp{ANN}, the loss function is used to adjust the weights of the \acs{ANN} in order to minimise the error \cite{ref:engelbrecht:2007}.

There are many variants of gradient-based heuristics, however, they all fundamentally apply the same generic \acf{GD} framework that propagates the error signal backwards through the \acs{ANN}. This algorithm is known as \acf{BP}. \Acs{BP} was popularised by \citeauthor{ref:werbos:1994} \cite{ref:werbos:1994}.

@@

Alternative variants have been proposed that lead to better control over the convergence characteristics caused by \acs{SGD}. This research focusses on a number of these variants that include \Acs{Momentum} \cite{ref:qian:1999}, \Acf{NAG} \cite{ref:sutskever:2013}, \Acf{Adagrad} \cite{ref:duchi:2011}, \acs{Adadelta} \cite{ref:zeiler:2012}, \Acf{RMSProp} \cite{ref:hinton:2012} and \Acf{Adam} \cite{ref:kingma:2014}.

@@

The biggest difference between \acp{MH} and gradient-bases heuristics is that
\acp{MH} make use of meta-information obtained as a result of evaluating the \acs{FFNN} during training and is not limited to information about the search space \cite{ref:blum:2003}. Similar to gradient-based heuristics, a number of different meta-heuristics have been used to successfully train \acp{FFNN} \cite{ref:rakitianskaia:2012, ref:vanwyk:2014, ref:espinal:2011, ref:gupta:1999}. This research takes a particular interest in population-based \acp{MH} that have been used to successfully train \acp{FFNN}. These include \acp{PSO} \cite{ref:shi:1998}, \acf{DE} \cite{ref:price:2006} and \Acp{GA} \cite{ref:fraser:1957}.

>>>>

Gradient-based heuristics are optimisation techniques that make use of derivates obtained from evaluating the \acs{ANN} being trained. In the context of supervised learning, loss functions produce a scalar value $\epsilon$, that represents the error between the output of the \acs{ANN} and the desired output. When using \acf{GD} to train \acp{ANN}, the loss function is used to adjust the weights of the \acs{ANN} in order to minimise the error \cite{ref:engelbrecht:2007}.

There are many variants of gradient-based heuristics, however, they all fundamentally apply the same generic \acs{GD} framework that propagates the error signal backwards through the \acs{ANN}. This algorithm is known as \acf{BP}. \acs{BP} was popularised by \citeauthor{ref:werbos:1994} \cite{ref:werbos:1994}.

@@

Alternative variants have been proposed that lead to better control over the convergence characteristics caused by \acs{SGD}. This research focusses on a number of these variants that include \acf{Momentum} \cite{ref:qian:1999}, \acf{NAG} \cite{ref:sutskever:2013}, \acf{Adagrad} \cite{ref:duchi:2011}, \acs{Adadelta} \cite{ref:zeiler:2012}, \acf{RMSProp} \cite{ref:hinton:2012} and \acf{Adam} \cite{ref:kingma:2014}.

@@

The biggest difference between \acp{MH} and gradient-bases heuristics is that
\acp{MH} make use of meta-information obtained as a result of evaluating the \acs{FFNN} during training and is not limited to information about the search space \cite{ref:blum:2003}. Similar to gradient-based heuristics, a number of different meta-heuristics have been used to successfully train \acp{FFNN} \cite{ref:rakitianskaia:2012, ref:vanwyk:2014, ref:espinal:2011, ref:gupta:1999}. This research takes a particular interest in population-based \acp{MH} that have been used to successfully train \acp{FFNN}. These include \acf{PSO} \cite{ref:shi:1998}, \acf{DE} \cite{ref:price:2006} and \acfp{GA} \cite{ref:fraser:1957}.

### 2023-02-04 article/sections/04-hyper-heuristics.tex
🚧 (article): WIP save point for revision

~\citeauthor{ref:burke:2010}~\cite{ref:burke:2010} define \acp{HH} as search methods or learning mechanism for selecting or generating heuristics to solve computational search problems.~\citeauthor{ref:burke:2003}~\cite{ref:burke:2003} mentions that a \acs{HH} is a high-level heuristic approach that, given a particular problem instance and a number of low-level heuristics, can select and apply an appropriate low-level heuristic at each decision point. \acp{HH} implement a form of meta-learning that is concerned with the selection of the best heuristic from a pool of heuristics to solve a given problem. It can be said that \acp{HH} are concerned with finding the best heuristic in \textit{heuristic space}, while the underlying low-level heuristics find solutions in the feasible \textit{search/solution space}.

\citeauthor{ref:burke:2010}~\cite{ref:burke:2010} proposed a modern classification scheme used to classify \acp{HH}. According to the proposed classification scheme, \acp{HH} are classified in two dimensions. These include the \textit{source of feedback} used during learning and the nature of the \textit{heuristic search space}. For the dimension that involves the source of feedback, \acp{HH} can be classified as either \textit{no learning}, \textit{online learning} or \textit{offline learning}. For the dimension that involves the nature of the \textit{heuristic search space}, \acp{HH} can be classified as either \textit{heuristic selection} or \textit{heuristic generation}. Further distinction is made between construction heuristics and perturbation heuristics.

In the general context of optimisation, many different types of \acp{HH} have been implemented and applied to many different problems. Some notable examples include \cite{ref:dowsland:2007, ref:burke:2010, ref:grobler:2012, ref:vanderstockt:2018}. Research on the application of \acp{HH} in the context of \acs{FFNN} training is still scarce.~\citeauthor{ref:nel:2021}~\cite{ref:nel:2021} provides some of the first research in this field, applying a \acs{HH} to \acs{FFNN} training.

>>>>

\citeauthor{ref:burke:2010} \cite{ref:burke:2010} define \acp{HH} as search methods or learning mechanism for selecting or generating heuristics to solve computational search problems. \citeauthor{ref:burke:2003} \cite{ref:burke:2003} mentions that a \acs{HH} is a high-level heuristic approach that, given a particular problem instance and a number of low-level heuristics, can select and apply an appropriate low-level heuristic at each decision point. \acp{HH} implement a form of \textit{meta-learning} that is concerned with the selection of the best heuristic from a pool of heuristics to solve a given problem. It can be said that \acp{HH} are concerned with finding the best heuristic in \textit{heuristic space}, while the underlying low-level heuristics find solutions in the feasible \textit{search/solution space}.

\citeauthor{ref:burke:2010} \cite{ref:burke:2010} proposed a modern classification scheme used to classify \acp{HH}. According to the proposed classification scheme, \acp{HH} are classified in two dimensions. These include the \textit{source of feedback} used during learning and the nature of the \textit{heuristic search space}. For the dimension that involves the source of feedback, \acp{HH} can be classified as either \textit{no learning}, \textit{online learning} or \textit{offline learning}. For the dimension that involves the nature of the \textit{heuristic search space}, \acp{HH} can be classified as either \textit{heuristic selection} or \textit{heuristic generation}. Further distinction is made between \textit{construction} of heuristics and \textit{perturbation} of heuristics.

In the general context of optimisation, many different types of \acp{HH} have been implemented and applied to many different problems. Some notable examples include \cite{ref:dowsland:2007, ref:burke:2010, ref:grobler:2012, ref:vanderstockt:2018}. Research on the application of \acp{HH} in the context of \acs{FFNN} training is still scarce. \citeauthor{ref:nel:2021} \cite{ref:nel:2021} provides some of the first research in this field, applying a \acs{HH} to \acs{FFNN} training.

### 2023-02-04 article/sections/05-probability.tex
🚧 (article): WIP save point for revision

Bayesian statistics describe the probability of an event in terms of some belief, based on previous knowledge of the event and the conditions under which the event happened~\cite{ref:hackenberger:2019}. Bayes' theorem expresses how a degree of belief, expressed as a probability, should rationally change to account for the availability of related evidence.

@@

Bayesian analysis utilises the the concept of conjugate priors. \citeauthor{ref:wackerly:2014}\cite{ref:wackerly:2014} state that conjugate priors are prior probability distributions that result in posterior distributions that are of the same functional form, $\mathcal{A}(v)$, as the prior, but with different parameter values. The conjugate prior to a Bernoulli probability distribution is the Beta probability distribution The conjugate prior to a Categorical and Multinomial probability distribution is the Dirichlet probability distribution~\cite{ref:wackerly:2014}.

>>>>

Bayesian statistics describe the probability of an event in terms of some belief, based on previous knowledge of the event and the conditions under which the event happened \cite{ref:hackenberger:2019}. Bayes' theorem expresses how a degree of belief, expressed as a probability, should rationally change to account for the availability of related evidence.

@@

Bayesian analysis utilises the the concept of conjugate priors. \citeauthor{ref:wackerly:2014}\cite{ref:wackerly:2014} state that conjugate priors are prior probability distributions that result in posterior distributions that are of the same functional form, $\mathcal{A}(v)$, as the prior, but with different parameter values. The conjugate prior to a Bernoulli probability distribution is the Beta probability distribution and the conjugate prior to a Categorical and Multinomial probability distribution is the Dirichlet probability distribution \cite{ref:wackerly:2014}.

### 2023-02-04 article/sections/06-bayesian-hyper-heuristics.tex
🚧 (article): WIP save point for revision

According to the classification scheme for \acp{HH} by~\citeauthor{ref:burke:2010}~\cite{ref:burke:2010}, the \Acs{BHH} is a population-based, meta-hyper-heuristic that utilises selection and perturbation of low-level heuristics in an online learning fashion. An illustration of the high-level architecture of the \Acs{BHH} is given in Appendix \ref{app:appendix_bhh_architecture} and the high level pseudo-code implementation of the \acs{BHH} is given in Appendix \ref{app:appendix_bhh_algorithm}. Discussions follow on the most important components of the \acs{BHH}.

@@

Generally speaking, the heuristic pool is a collection of low-level heuristics under consideration by the \acs{BHH}. The heuristic pool contains the set of low-level heuristics that, together with their performance information, make up the \text{heuristic space}. Importantly, The heuristic pool must consist of a diverse set of low-level heuristics with varying capabilities. This research takes an interest in including both gradient-based heuristics as well as \acp{MH} in the heuristic pool.

@@

The concept of proxies arise from the sparsity of state as maintained by different heuristics. Since heuristics maintain (possibly) different states, there is an uncertainty of state transition when switching between heuristics. A solution to state indifference is to proxy heuristic state update operations. State is then maintained in two parts: primary and proxied state. Primary state refers to the state that is originally maintained by a heuristic. While proxied state refers to the state that is not directly maintained by the heuristic, but can be updated by outsourcing the required state update operation to another heuristic. The \acs{BHH} thus incorporates a mapping of proxied state update operations.

@@

Entities represent candidate solutions to the model's trainable parameters (weights) and other heuristic-specific state parameters. These state parameters are referred to as \textit{local} state. Entities are treated as physical \textit{particles} in a hyper-dimensional search environment. Entities model concepts from physics. For example, the candidate solution is represented as the entity's position, and velocity and acceleration is analogous to the gradient and momentum of the entity respectfully. Examples of entity state parameters, as derived from various low-level heuristics, include entity position, velocity, gradient, position delta, first and second moments of the gradient, the loss, personal best positions and losses, and so on. The entity state parameters are updated by the associated heuristic.

@@

The credit assignment strategy is a mechanism that assigns a discrete credit indicator to heuristics that perform well, based on their performance metrics such as loss. The credit assignment strategy implements the ``move acceptance'' process as proposed by~\citeauthor{ref:ozcan:2006}~\cite{ref:ozcan:2006,ref:ozcan:2008} and addresses the credit assignment problem as discussed by~\citeauthor{ref:burke:2010}~\cite{ref:burke:2010}. A good credit assignment strategy will correctly allocate credit to the appropriate heuristic-entity combination. This research  implements the following credit assignment strategies to choose from: \textit{ibest} (iteration best), \textit{pbest} (personal best), \textit{gbest} (global best), \textit{rbest} (replay window best), and \textit{symmetric}, where credit is assigned to all entity-heuristic combinations, regardless of their performance.

@@

The predictive model thus models the \textit{proportional},probability of the event (selection of) heuristic $\boldsymbol{H}$, given allocation to entity $\boldsymbol{E}$ and credit $\boldsymbol{C}$, parameterised by sampled $\boldsymbol{\theta}$, $\boldsymbol{\phi}$ and $\boldsymbol{\psi}$. These parameters are in turn parameterised by concentrations $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$, $\gamma_{1}$ and $\gamma_{0}$, which denote the prior beliefs of the \acs{BHH}.

@@

Generally, there are two different techniques that are used to train Naïve Bayes classifiers. The frequentist approach implements \acs{MLE} and the Bayesian approach implements \acs{MAP}.

@@

It can be said that the \acs{BHH} implements a Gaussian process~\cite{ref:gortler:2019}. Since the reselection of heuristics happens at regular intervals, the outcome of a selection in one iteration may influence the outcome of another in the next iteration, making the implementation of the \acs{BHH} a \acs{HMM}~\cite{ref:rabiner:1986}.

>>>>

According to the classification scheme for \acp{HH} by \citeauthor{ref:burke:2010} \cite{ref:burke:2010}, the \acs{BHH} is a population-based, meta-hyper-heuristic that utilises selection and perturbation of low-level heuristics in an online learning fashion.

The reader is referred to \cite{ref:schreuder:2022} for an illustration of the high-level architecture of the \acs{BHH} as well as the high level pseudo-code implementation of the \acs{BHH}. Discussions follow on the most important components of the \acs{BHH}.

@@

Generally speaking, the heuristic pool is a collection of low-level heuristics under consideration by the \acs{BHH}. The heuristic pool contains the set of low-level heuristics that, together with their performance information, make up the \text{heuristic space}. Importantly, the heuristic pool must consist of a diverse set of low-level heuristics with varying capabilities. This research takes an interest in including both gradient-based heuristics as well as \acp{MH} in the heuristic pool. This approach is referred to as a \textit{multi-method} approach.

@@

The concept of proxies arise from the sparsity of state as maintained by different heuristics. Since heuristics maintain (possibly) different states, there is an uncertainty of state transition when switching between heuristics. A solution to state indifference is to \textit{proxy} heuristic state update operations. State is then maintained in two parts: primary and proxied state. Primary state refers to the state that is originally maintained by a heuristic. While proxied state refers to the state that is not directly maintained by the heuristic, but can be updated by outsourcing the required state update operation to another heuristic. The \acs{BHH} thus incorporates a mapping of proxied state update operations. Similar to before, the reader is referred to \cite{ref:schreuder:2022} for details on the mapped proxied state update operations.

@@

Entities represent candidate solutions to the model's trainable parameters (weights) and other heuristic-specific state parameters. These state parameters are referred to as \textit{local} state. Entities are treated as physical \textit{particles} in a hyper-dimensional search environment. Entities model concepts from physics. For example, the candidate solution is represented as the entity's position, and velocity and acceleration is analogous to the gradient and momentum of the entity respectfully. Examples of entity state parameters, as derived from various low-level heuristics, include entity position, velocity, gradient, position delta, first and second moments of the gradient, the loss, personal best positions, losses, and so on. The entity state parameters are updated by the associated heuristic.

@@

The credit assignment strategy is a mechanism that assigns a discrete credit indicator to heuristics that perform well, based on their performance metrics such as loss. The credit assignment strategy implements a component of the ``move acceptance'' process as proposed by \citeauthor{ref:ozcan:2006} \cite{ref:ozcan:2006,ref:ozcan:2008} and addresses the credit assignment problem as discussed by \citeauthor{ref:burke:2010} \cite{ref:burke:2010}. A good credit assignment strategy will correctly allocate credit to the appropriate heuristic-entity combination. This research implements the following credit assignment strategies to choose from: \textit{ibest} (iteration best), \textit{pbest} (personal best), \textit{gbest} (global best), \textit{rbest} (replay window best), and \textit{symmetric}, where credit is assigned to all entity-heuristic combinations, regardless of their performance.

@@

The predictive model thus models the \textit{proportional} probability of the event (selection of) heuristic $\boldsymbol{H}$, given allocation to entity $\boldsymbol{E}$ and credit requirement $\boldsymbol{C}$, parameterised by sampled $\boldsymbol{\theta} \vert \boldsymbol{\alpha} \sim Dir(\boldsymbol{\alpha}; K)$, $\boldsymbol{\phi} \vert \boldsymbol{\beta} \sim Dir(\boldsymbol{\beta}; K)^{J}$ and $\boldsymbol{\psi} \vert \gamma_{1}, \gamma_{0}  \sim Beta(\gamma_{1}, \gamma_{0})$. In the aforementioned, $K$ is the heuristic pool size, $J$ is the entity pool size and the indices $1$ and $0$ refers to success and failure parameters respectively. The concentration parameters $\boldsymbol{\alpha}$, $\boldsymbol{\beta}$, $\gamma_{1}$ and $\gamma_{0}$ denote the prior beliefs of the \acs{BHH}. Similar to before, the reader is referred to \cite{ref:schreuder:2022} for the mathematical derivation of the selection mechanism.

@@

Generally, there are two different techniques that are used to train Naïve Bayes classifiers. The frequentist approach implements \acf{MLE} and the Bayesian approach implements \acf{MAP}.

@@

It can be said that the \acs{BHH} implements a Gaussian process \cite{ref:gortler:2019}. Since the reselection of heuristics happens at regular intervals, the outcome of a selection in one iteration may influence the outcome of another in the next iteration, making the implementation of the \acs{BHH} a \acs{HMM} \cite{ref:rabiner:1986}.

### 2023-02-04 article/sections/07-methodology.tex
🚧 (article): WIP save point for revision

In the context of training \acp{FFNN}, the underlying models are trained across a number of datasets. All the datasets used in the empirical process originate from the UCI Machine Learning Repository~\cite{ref:uci:2022}. Datasets are grouped by problem type and include seven classification and seven regression datasets. The details around the datasets used can be found in Tables \ref{tab:methodology:datasets:classification} and \ref{tab:methodology:datasets:regression}. Each dataset is split into a training set comprising of 80\% of the data, and a test set comprising of 20\% of the data. The test set is used as a validation set during training.

@@

All models that are trained in the empirical process follow implementations of shallow \acp{FFNN} with only one hidden layer. The number of hidden units used were determined empirically. Weights are initialised by means of Glorot uniform sampling~\cite{ref:glorot:2010}. The models and their configuration, as it is used for each dataset, are given in Table \ref{tab:methodology:models:configurations}.

@@

The Shapiro-Wilk test for normality~\cite{ref:shapiro:1965} ($\alpha$ = 0.001) is used to determine if the results are normally distributed. The Levene's test for equality of variance~\cite{ref:levene:1961} ($\alpha$ = 0.001) is used. For experiments with three or more configurations, the \acs{ANOVA} statistical test~\cite{ref:fisher:1921} ($\alpha$ = 0.001) is used. The Kruskal-Wallis ranked non-parametric test~\cite{ref:kruskal:1952} for statistical significance ($\alpha$ = 0.001) is used for cases where data is not normally distributed. Finally, a post-hoc Tukey honest significant difference test~\cite{ref:tukey:1949} ($\alpha$ = 0.001) is used from which significant ranking is retrieved. Descriptive and critical difference plots are then retrieved from these results to provide visual aid.

>>>>

In the context of training \acp{FFNN}, the underlying models are trained across a number of datasets. All the datasets used in the empirical process originate from the UCI Machine Learning Repository \cite{ref:uci:2022}. Datasets are grouped by problem type and include seven classification and seven regression datasets. The details around the datasets used can be found in Tables \ref{tab:methodology:datasets:classification} and \ref{tab:methodology:datasets:regression}. Each dataset is split into a training set comprising of 80\% of the data, and a test set comprising of 20\% of the data. The test set is used as a validation set during training.

@@

All models that are trained in the empirical process follow implementations of shallow \acp{FFNN} with only one hidden layer. The number of hidden units used were determined empirically. Weights are initialised by means of Glorot uniform sampling \cite{ref:glorot:2010}. The models and their configuration, as it is used for each dataset, are given in Table \ref{tab:methodology:models:configurations}.

@@

The Shapiro-Wilk test for normality \cite{ref:shapiro:1965} ($\alpha$ = 0.001) is used to determine if the results are normally distributed. The Levene's test for equality of variance \cite{ref:levene:1961} ($\alpha$ = 0.001) is used. For experiments with three or more configurations, the \acs{ANOVA} statistical test \cite{ref:fisher:1921} ($\alpha$ = 0.001) is used. The Kruskal-Wallis ranked non-parametric test \cite{ref:kruskal:1952} for statistical significance ($\alpha$ = 0.001) is used for cases where data is not normally distributed. Finally, a post-hoc Tukey honest significant difference test \cite{ref:tukey:1949} ($\alpha$ = 0.001) is used from which significant ranking is retrieved. Descriptive and critical difference plots are then retrieved from these results to provide visual aid.
